Word Learning as Network Growth: A Cross-linguistic Analysis

### 2019-02-05 journal/ms.Rmd
modif journal

```{r load_packages, include = FALSE}
library(papaja)
library(readr)
library(tidyr)
library(ggplot2)
library(cowplot)
library(knitr)
library(boot)
library(dplyr)
library(langcog)
library(ggthemes)
library(broom)
library(HDInterval)
library(kableExtra)
#library("knitcitations")
#cleanbib()
#options("citation_format" = "pandoc")
```

@@

What factors shape vocabulary learning over the course of early childhood? To investigate this question, scientists have adopted multiple research strategies, from conducting controlled laboratory experiments [e.g. @markman90] to analyzing dense corpora capturing language learning in context [e.g., @roy2015]. One strategy consists in documenting the timeline of words' acquisition, and studying the properties that make words easy or hard to learn. For example, within a lexical category, words that are more frequent in child-directed speech are acquired earlier [@goodman2008]. Other factors include word length and the mean length of utterances in which the word occurs [e.g., @braginsky2016; @swingley2018].

Besides these word-level properties, the structure of the lexicon (that is, how words relate to one another) also predicts the Age of Acquisition (AoA) of words.  The lexical structure can be characterized in terms of a network where each node represents a word in the vocabulary, and each link between two nodes represents a relationship between the corresponding pair of words [e.g., @collins1975]. Previous studies have investigated early vocabulary structure by constructing networks using  a variety of word-word relations including shared semantic features [@mcrae2005], target-cue relationships in free association norms [@nelson1998], co-occurrence in child directed speech [@macwhinney2014], and phonological relatedness [@vitevitch2008]. These studies have generally found that children tend to produce words that have higher neighborhood density (i.e., high connectivity in the network) earlier, both at the phonological and the semantic level [@storkel2009; @hills2009; @hills2010; @stella2017; @carlson2014].

@@

These two proposals represent two divergent ideas about the role of lexical networks in acquisition. On the INT proposal, learning is driven by known words with high connectivity to other known words (Figure 1, left). Thus, the network structure is a causal factor in early word learning, that is, children rely on the organization of their past knowledge to determine future learning. For example, having a rich and organized knowledge about the domain of dinosaurs would facilitate the acquisition of new dinosaur-related words [@chi1983]. In contrast, on the EXT approach, learning is driven by the connectivity of words that are not known yet (Figure 1, right). Thus, the relevant network structure is not internally represented by children, and the observed connectivity effect might be an epiphenomenon of some properties of the linguistic input. For example, highly connected words in the input could be more easily learned because of their contextual diversity [e.g., @smith2008] or because caregivers emphasize such words in child-directed speech [@macwhinney2014].

@@

@hills2009 investigated the growth of lexico-semantic networks in toddlers and found that growth did not proceed according to INT as was originally hypothesised by @steyvers2005, but rather according to EXT.\footnote{Besides INT and EXT, the authors tested a third mechanism (called the lure of associates) which resembles EXT in that it is driven by the connectivity of external nodes, except that this connectivity is computed with respect to words that are known. However, EXT is the externally-driven scenario that best predicted the data in this previous work.} This is a profound finding because it suggests that learning in the early stages is mostly driven by properties of the external input, regardless of how past knowledge is organized. However, this work has been done in the special case of networks that are based on 1) semantic associations, 2) production-based vocabularies, and 3) data from English-learning children, only. The extent to which this result depends on the domain (e.g., semantic vs. phonological connectivity), the vocabulary measure (production vs. comprehension) and culture/language is still unclear. In this work, we test the generality of the finding along these three dimensions.

@@

We test the growth scenarios using parent reports of children vocabularies and their normative age of acquisition [@fenson94].  Children may vary in their individual learning trajectories, but the aggregate data leads to an average learning pattern which is highly consistent. It is this normative trajectory that we model in the current study, following in this regard the steps of previous research using similar datasets [@storkel2009; @stella2017; @hills2009; @hills2010; @goodman2008; @braginsky2016]

@@

plot_correlation_prod <- ggplot(subset(data_long, measure == "produces"), aes(x=value, y=aoa))+
  facet_grid(predictor ~ language)+#, scales = "free") +
  geom_jitter(#aes(colour = lexical_class),
    colour = solarized_palette(1),
    size = 0.5, alpha = 0.5) +
  geom_abline(slope = -1)+
  coord_cartesian(xlim=c(-1,5))+
  #scale_x_continuous(limits=c(-2,5))+
  scale_y_continuous(breaks =c(15,25,35))+
  geom_smooth(method = "lm", colour = "grey3", se=FALSE)+
  scale_colour_solarized(name = "") +
  theme_few()+
  #theme_bw()+
  theme(aspect.ratio = 0.7, 
        plot.margin=grid::unit(c(0,0,0,0), "mm")
        )+
  geom_text(data=subset(correlations, measure=="produces"), aes(label=paste("r=", cor, sep="")), x=3.5, y=33, size=2, fontface = "bold")+
  xlab("degree z-score") +ylab("Age")

plot_correlation_comp <- ggplot(subset(data_long, measure == "understands" & aoa < 35), aes(x=value, y=aoa))+
  facet_grid(predictor ~ language)+#, scales = "free") +
  geom_jitter(#aes(colour = lexical_class),
    colour = solarized_palette(1),
    size = 0.5, alpha = 0.5) +
  geom_abline(slope = -1)+
  coord_cartesian(xlim=c(-1,5))+
  #scale_x_continuous(limits=c(-2,5))+
  scale_y_continuous(breaks =c(15,25,35))+
  geom_smooth(method = "lm", colour = "grey3", se=FALSE)+
  scale_colour_solarized(name = "") +
  theme_few()+
  #theme_bw()+
  theme(aspect.ratio = 0.7, 
        plot.margin=grid::unit(c(0,0,0,0), "mm")
        )+
  geom_text(data=subset(correlations, measure=="understands"), aes(label=paste("r=", cor, sep="")), x=3.5, y=25, size=2, fontface = "bold")+
  xlab("degree z-score") +ylab("Age")

@@

```{r corrAll, out.width = "\\textwidth", fig.pos = "!h", fig.cap = "Age of production (A) and comprehension (B) in the global network as predicted by the degree (i.e., connectivity) in this network. Results are shown in each language for phonological and semantic networks. Each point is a word, with lines indicating linear model fits, and numbers indicating the Pearson correlation coefficients.", fig.width=8, fig.height=4}

plot_grid(plot_correlation_prod , NULL, plot_correlation_comp , labels = c("A", "", "B"), ncol = 1, align = "v", rel_heights = c(1.3, 0.1, 1.3))

@@

Figure \@ref(fig:corrAll) shows how the age of production and comprehension for each word varies as a function of its degree (or indegree for the semantic networks) as well as the correlation values. For ease of visual comparison, the predictor (i.e., the degree) was centered and scaled across languages. The plots show, overall, a negative correlation between the month of acquisition and the degree.  In production data, the average correlation across languages was `r ps_ave` (SD=`r ps_sd`) for the semantic networks and `r pp_ave` (SD=`r pp_sd`) for the phonological networks.  In comprehension data, the average correlation was `r cs_ave` (SD=`r cs_sd`) for the semantic networks and `r cp_ave` (SD=`r cp_sd`) for the phonological networks. These results indicate that nouns with higher degrees are generally learned earlier, thus replicating previous findings in English (e.g., Storkel 2004, 2009; Hills et al. 2009) and extending these findings to 10 different languages, generally, in both production- and comprehension-based vocabularies.

@@

plot_degree <- ggplot(data = degree_dist,  aes(x=x, y=y, col=dimension))+
  facet_grid(data ~ language)+#, scales = "free") +
  geom_point(#aes(colour = lexical_class),
    #colour = solarized_palette(1),
    size = 0.5, alpha = 0.5) +
  scale_y_log10() + scale_x_log10() +
  theme(aspect.ratio = 1) +
  theme_few()+
  theme(aspect.ratio = 0.7)+
  xlab("Degree") +ylab("Prob.")

ggsave("figs/plot_degree.pdf", plot = plot_degree, width = 10, height = 3)
```

```{r degreeDist, out.width = "\\textwidth", fig.pos = "!h", fig.cap = "Log-log plot of the cumulative degree distribution function for the global phonological and semantic networks across languages. The figure shows the results for both production and comprehension data. A perfect power-law distribution should appear as a straight line in this graph.",  fig.width=10, fig.height=2}

#knitr::include_graphics("figs/plot_degree.pdf")
plot_degree
```

@@

For word length, we counted the number of phonemes in our generated IPA transcription. For word frequency, we used the frequency estimates from @braginsky2016 where unigram counts were derived based on CHILDES corpora in each language [@macwhinney2014]. For each word, counts included words that shared the same stem  (e.g., "cats" counts as "cat"), or words that were synonymous (e.g. "father" counts as "daddy"). Although these frequency counts use transcripts from independent sets of children, they are based on large samples, and this allows us to average out possible differences between children and the specificities of their input [see @goodman2008].

>>>>

```{r load_packages, include = FALSE}
library(papaja)
library(readr)
library(tidyr)
library(ggplot2)
library(cowplot)
library(ggpubr)
library(knitr)
library(boot)
library(dplyr)
library(gridExtra)
library(langcog)
library(ggthemes)
library(broom)
library(HDInterval)
library(kableExtra)
#library("knitcitations")
#cleanbib()
#options("citation_format" = "pandoc")
```

@@

What factors shape vocabulary learning over the course of early childhood? To investigate this question, scientists have adopted multiple research strategies, from conducting controlled laboratory experiments [e.g. @markman90] to analyzing dense corpora capturing language learning in context [e.g., @roy2015]. One prominent strategy consists in documenting the timeline of words' acquisition, and studying the properties that make words easy or hard to learn [e.g., @goodman2008; @huttenlocher1991]. For example, @goodman2008 found that, within a lexical category (e.g., nouns), higher parental frequency is associated with earlier learning. Researchers have studied the role of several other factors such as word length and the mean length of utterances in which the word occurs [e.g., @braginsky2016; @swingley2018].

Besides word-level properties, the structure of the lexicon (that is, how words relate to one another) was also found to predict the Age of Acquisition (AoA) of words. The lexical structure can be characterized in terms of a network where each node represents a word in the vocabulary, and each link between two nodes represents a relationship between the corresponding pair of words [e.g., @collins1975; @luce1998]. Previous studies have investigated early vocabulary structure by constructing networks using  a variety of word-word relations including shared semantic features [@mcrae2005], target-cue relationships in free association norms [@nelson1998], co-occurrence in child directed speech [@macwhinney2014], and phonological relatedness [@vitevitch2008]. These studies have generally found that children tend to produce words that have higher neighborhood density (i.e., high connectivity in the network) earlier, both at the phonological and the semantic level [@storkel2009; @hills2009; @hills2010; @stella2017; @carlson2014].

@@

<!-- For example, having a rich and organized knowledge about the domain of dinosaurs would facilitate the acquisition of new dinosaur-related words [@chi1983]. -->
These two proposals represent two divergent ideas about the role of lexical networks in acquisition. On the INT proposal, learning is driven by known words with high connectivity to other known words (Figure 1, left). Thus, the network structure is a causal factor in early word learning, that is, children rely on the organization of their past knowledge to determine future learning [@storkel2009; @altvater2013; @chi1983; @borovsky2016]. In contrast, on the EXT approach, learning is driven by the connectivity of words that are not known yet (Figure 1, right). Thus, the relevant network structure is not internally represented by children, and the observed connectivity effect might be an epiphenomenon of some properties of the linguistic input. For example, highly connected words in the input could be more easily learned because of their contextual diversity, allowing for an easier meaning disambiguation [@smith2008; @mcmurray2012; @yurovsky2015]. Another reason could be that these words are emphasized by the cargivers in their child-directed speech [@huttenlocher1991; @hoff2002; @clark2007].

@@

@hills2009 investigated the growth of lexico-semantic networks in toddlers and found that growth did not proceed according to INT as was originally hypothesised by @steyvers2005, but rather according to EXT.\footnote{Besides INT and EXT, the authors tested a third mechanism (called the lure of associates) which resembles EXT in that it is driven by the connectivity of external nodes, except that this connectivity is computed with respect to words that are known. However, EXT is the externally-driven scenario that best predicted the data in this previous work.} This is an important finding because it suggests that learning in the early stages is mostly driven by properties of the external input, regardless of how past knowledge is organized. However, this work has been done in the special case of networks that are based on 1) semantic associations, 2) production-based vocabularies, and 3) data from English-learning children, only. The extent to which this result depends on the domain (e.g., semantic vs. phonological connectivity), the vocabulary measure (production vs. comprehension) and culture/language is still unclear. In this work, we test the generality of the finding along these three dimensions.

@@

We adopted the following research strategy. We used a cross-linguistic dataset of children's vocabulary development [@frank2017; @fenson94], and we studied the timeline of word learning using the normative/average age of acquisition (i.e., the age at which at least 50% of children know a given words). Our choice of studying the normative learning trajectory rather than the individual trajectories was motivated by the nature of the dataset used---which is primarily based on cross-sectional studies. Children may vary in their individual learning trajectories, but the aggregate data provide highly robust measures of the average learning patterns [@fenson94]. The use of such measures has lead to important insights on the mechanisms of word learning [@storkel2009; @stella2017; @hills2009; @hills2010; @goodman2008; @braginsky2016].

@@

plot_correlation_prod_1 <- ggplot(subset(data_long, measure == "produces" & data_long$language %in% c('Croatian', 'Danish', 'English', 'French','Italian')), 
                                  aes(x=value, y=aoa))+
  facet_grid(language~ predictor)+#, scales = "free") +
  geom_jitter(#aes(colour = lexical_class),
    colour = solarized_palette(1),
    size = 0.5, alpha = 0.5) +
  geom_abline(slope = -1)+
  coord_cartesian(xlim=c(-1,5))+
  #scale_x_continuous(limits=c(-2,5))+
  scale_y_continuous(breaks =c(15,25,35))+
  geom_smooth(method = "lm", colour = "grey3", se=FALSE)+
  scale_colour_solarized(name = "") +
  theme_few()+
  #theme_bw()+
  theme(aspect.ratio = 0.7, 
        plot.margin=grid::unit(c(0,0,0,0), "mm")
        )+
  geom_text(data=subset(correlations, measure=="produces" & correlations$language %in% c('Croatian', 'Danish', 'English', 'French','Italian')), aes(label=paste("r=", cor, sep="")), x=3.5, y=33, size=2, fontface = "bold")+
  xlab("degree z-score") +ylab("Age")

plot_correlation_prod_2 <- ggplot(subset(data_long, measure == "produces" & data_long$language %in% c('Norwegian', 'Russian', 'Spanish', 'Swedish','Turkish')), aes(x=value, y=aoa))+
  facet_grid(language~ predictor)+#, scales = "free") +
  geom_jitter(#aes(colour = lexical_class),
    colour = solarized_palette(1),
    size = 0.5, alpha = 0.5) +
  geom_abline(slope = -1)+
  coord_cartesian(xlim=c(-1,5))+
  #scale_x_continuous(limits=c(-2,5))+
  scale_y_continuous(breaks =c(15,25,35))+
  geom_smooth(method = "lm", colour = "grey3", se=FALSE)+
  scale_colour_solarized(name = "") +
  theme_few()+
  #theme_bw()+
  theme(aspect.ratio = 0.7, 
        plot.margin=grid::unit(c(0,0,0,0), "mm")
        )+
  geom_text(data=subset(correlations, measure=="produces" & correlations$language %in% c('Norwegian', 'Russian', 'Spanish', 'Swedish','Turkish')), aes(label=paste("r=", cor, sep="")), x=3.5, y=33, size=2, fontface = "bold")+
  xlab("degree z-score") +ylab("Age")

corr_prod <- cowplot::plot_grid(
  plot_correlation_prod_1 ,
  plot_correlation_prod_2 ,
  nrow = 1#, 
  #align = 'vh',
  #hjust = -1
)

#legend <- cowplot::get_legend(plot_degree_1)

#plot_degree <- cowplot::plot_grid(prow, legend, nrow = 1, rel_widths = c(4, .5))

plot_correlation_comp_1 <- ggplot(subset(data_long, measure == "understands" & aoa < 35 & data_long$language %in% c('Croatian', 'Danish', 'English', 'French','Italian')), aes(x=value, y=aoa))+
  facet_grid(language~ predictor)+#, scales = "free") +
  geom_jitter(#aes(colour = lexical_class),
    colour = solarized_palette(1),
    size = 0.5, alpha = 0.5) +
  geom_abline(slope = -1)+
  coord_cartesian(xlim=c(-1,5))+
  #scale_x_continuous(limits=c(-2,5))+
  scale_y_continuous(breaks =c(15,25,35))+
  geom_smooth(method = "lm", colour = "grey3", se=FALSE)+
  scale_colour_solarized(name = "") +
  theme_few()+
  #theme_bw()+
  theme(aspect.ratio = 0.7, 
        plot.margin=grid::unit(c(0,0,0,0), "mm")
        )+
  geom_text(data=subset(correlations, measure=="understands" & correlations$language %in% c('Croatian', 'Danish', 'English', 'French','Italian')), aes(label=paste("r=", cor, sep="")), x=3.5, y=25, size=2, fontface = "bold")+
  xlab("degree z-score") +ylab("Age")

plot_correlation_comp_2 <- ggplot(subset(data_long, measure == "understands" & aoa < 35 & data_long$language %in% c('Norwegian', 'Russian', 'Spanish', 'Swedish','Turkish')), aes(x=value, y=aoa))+
  facet_grid(language~ predictor)+#, scales = "free") +
  geom_jitter(#aes(colour = lexical_class),
    colour = solarized_palette(1),
    size = 0.5, alpha = 0.5) +
  geom_abline(slope = -1)+
  coord_cartesian(xlim=c(-1,5))+
  #scale_x_continuous(limits=c(-2,5))+
  scale_y_continuous(breaks =c(15,25,35))+
  geom_smooth(method = "lm", colour = "grey3", se=FALSE)+
  scale_colour_solarized(name = "") +
  theme_few()+
  #theme_bw()+
  theme(aspect.ratio = 0.7, 
        plot.margin=grid::unit(c(0,0,0,0), "mm")
        )+
  geom_text(data=subset(correlations, measure=="understands" & correlations$language %in% c('Norwegian', 'Russian', 'Spanish', 'Swedish','Turkish')), aes(label=paste("r=", cor, sep="")), x=3.5, y=25, size=2, fontface = "bold")+
  xlab("degree z-score") +ylab("Age")

corr_comp <- cowplot::plot_grid(
  plot_correlation_comp_1 ,
  plot_correlation_comp_2 ,
  nrow = 1#, 
  #align = 'vh',
  #hjust = -1
)

@@

```{r corrProd, out.width = "\\textwidth", fig.pos = "!h", fig.height=5, fig.cap = "Age of production in the global network as predicted by the degree (i.e., connectivity) in this network. Results are shown in each language for phonological and semantic networks. Each point is a word, with lines indicating linear model fits, and numbers indicating the Pearson correlation coefficients."}

corr_prod
#plot_grid(plot_correlation_prod , NULL, plot_correlation_comp , labels = c("A", "", "B"), ncol = 1, align = "v", rel_heights = c(1.3, 0.1, 1.3))

```{r corrComp, out.width = "\\textwidth", fig.pos = "!h", fig.height=5, fig.cap = "Age of comprehension in the global network as predicted by the degree (i.e., connectivity) in this network. Results are shown in each language for phonological and semantic networks. Each point is a word, with lines indicating linear model fits, and numbers indicating the Pearson correlation coefficients."}
corr_comp
#plot_grid(plot_correlation_prod , NULL, plot_correlation_comp , labels = c("A", "", "B"), ncol = 1, align = "v", rel_heights = c(1.3, 0.1, 1.3))

@@

Figures \@ref(fig:corrProd) and \@ref(fig:corrComp) show how the age of production and comprehension, respectively, correlates with the degree (or indegree for the semantic networks). For ease of visual comparison, the predictor (i.e., the degree) was centered and scaled across languages. The plots show, overall, a negative correlation between the month of acquisition and the degree.  In production data, the average correlation across languages was `r ps_ave` (SD=`r ps_sd`) for the semantic networks and `r pp_ave` (SD=`r pp_sd`) for the phonological networks.  In comprehension data, the average correlation was `r cs_ave` (SD=`r cs_sd`) for the semantic networks and `r cp_ave` (SD=`r cp_sd`) for the phonological networks. These results indicate that nouns with higher degrees are generally learned earlier, thus replicating previous findings in English (e.g., Storkel 2004, 2009; Hills et al. 2009) and extending these findings to 10 different languages, generally, in both production- and comprehension-based vocabularies.

@@

plot_degree_1 <- ggplot(data = subset(degree_dist, degree_dist$language %in% c('Croatian', 'Danish', 'English', 'French','Italian')),  
                        aes(x=x, y=y, col=dimension))+
  facet_grid( language ~ data)+#, scales = "free") +
  geom_point(#aes(colour = lexical_class),
    #colour = solarized_palette(1),
    size = 0.5, alpha = 0.5) +
  scale_y_log10() + scale_x_log10() +
  theme_few()+
  theme(legend.title = element_text(size=11),
      legend.text=element_text(size=11),
      axis.text = element_text(size = 8),
      aspect.ratio = 0.7
      )+
  xlab("Degree") +ylab("")

plot_degree_2 <- ggplot(data = subset(degree_dist, degree_dist$language %in% c('Norwegian', 'Russian', 'Spanish', 'Swedish','Turkish')),  
                        aes(x=x, y=y, col=dimension))+
  facet_grid( language ~ data)+#, scales = "free") +
  geom_point(#aes(colour = lexical_class),
    #colour = solarized_palette(1),
    size = 0.5, alpha = 0.5) +
  scale_y_log10() + scale_x_log10() +
  theme_few()+
  theme(legend.title = element_text(size=11),
      legend.text=element_text(size=11),
      axis.text = element_text(size = 8),
      aspect.ratio = 0.7
      )+
  xlab("Degree") +ylab("Prob.")

prow <- cowplot::plot_grid(
  plot_degree_1 + theme(legend.position = "none"),
  plot_degree_2 + theme(legend.position = "none"),
  nrow = 1, 
  align = 'vh',
  hjust = -1
)
legend <- cowplot::get_legend(plot_degree_1)

plot_degree <- cowplot::plot_grid(prow, legend, nrow = 1, rel_widths = c(4, .5))

#plot_degree_2
#ggarrange(plot_degree_1, plot_degree_2, common.legend = TRUE, legend="bottom")
#ggsave("figs/plot_degree.png", plot = plot_degree, width = 4, height = 8)
#legend <- get_legend(plot_degree_1)

#plot_degree <- grid.arrange(
#  plot_degree_1 + theme(legend.position="none"),
#  plot_degree_2+  theme(legend.position="none"),
# legend,
#  ncol=3)

```{r degreeDist, out.width = "\\textwidth", fig.pos = "!h", fig.height=5, fig.cap = "Log-log plot of the cumulative degree distribution function for the global phonological and semantic networks across languages. The figure shows the results for both production and comprehension data. A perfect power-law distribution should appear as a straight line in this graph."}
#knitr::include_graphics("figs/plot_degree.pdf")
plot_degree

#img <- png::readPNG("figs/plot_degree.png")
#grid::grid.raster(img)
```

@@

For word length, we counted the number of phonemes in our generated IPA transcription. For word frequency, we used the frequency estimates from @braginsky2016 where unigram counts were derived based on CHILDES corpora in each language [@macwhinney2014]. For each word, counts included words that shared the same stem  (e.g., "cats" counts as "cat"), or words that were synonymous (e.g. "father" counts as "daddy"). Although these frequency counts use transcripts from independent sets of children, they are based on large samples, and this allows us to average out possible differences between children and the specificities of their input [see @goodman2008 for a similar research strategy].
